\documentclass[11pt,a4paper]{article}

\usepackage[top=1.5cm, bottom=1.2cm, left=1.8cm, right=1.8cm]{geometry}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{parskip}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{mdframed}

\definecolor{nured}{RGB}{200,16,46}
\definecolor{notebg}{RGB}{255,248,220}
\definecolor{noteborder}{RGB}{200,160,0}

%% Header / Footer
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0.4pt}
\lhead{\textcolor{nured}{\textbf{EECE5155}} --- Wireless Sensor Networks \& IoT \textbar\ Spring 2026}
\rhead{\textbf{MVP Hardware Specification}}
\rfoot{\small Due: \textbf{Wednesday March 19, end of day} --- submit PDF on Canvas}

%% Section formatting: compact red-underlined headings
\titleformat{\section}{\small\bfseries\color{nured}}{}{0em}{}[{\color{nured}\vspace{-5pt}\rule{\linewidth}{0.4pt}}]
\titlespacing{\section}{0pt}{5pt}{2pt}

% -----------------------------------------------------------------------
\begin{document}
\setlength{\parskip}{2pt}

% ========================= TITLE BLOCK =================================
\begin{center}
  {\large\bfseries MVP Hardware Specification}\\[3pt]
  {\small Northeastern University \textbar\ EECE5155: Wireless Sensor Networks and the IoT \textbar\ Spring 2026}\\[2pt]
  {\small\textit{Individual submission. One page maximum.
  Raspberry Pi and Arduino boards are \textbf{excluded} from the \$50 budget
  (they are provided). Everything else you need must appear in the BOM below.}}
\end{center}

\vspace{1pt}\noindent\rule{\linewidth}{0.6pt}\vspace{3pt}

% ========================= STUDENT INFO ================================
\section{Student \& Project}

\begin{tabularx}{\linewidth}{@{}lX@{\hspace{10mm}}lX@{}}
  \textbf{Name:}          & \hrulefill &
  \textbf{Project title:} & \hrulefill \\[5pt]
  \textbf{NUID:}          & \hrulefill &
  \textbf{MCU / SBC:}     & \hrulefill \\[2pt]
\end{tabularx}

% ========================= PES REQUIREMENTS ============================
\section{Requirements from your PES (copy the 3 most hardware-relevant)}

\begin{tabularx}{\linewidth}{@{}p{0.06\linewidth} X p{0.32\linewidth}@{}}
  \toprule
  \textbf{\#} & \textbf{Requirement (from PES)} & \textbf{Hardware implication} \\
  \midrule
  R1 & & \\[6pt]
  R2 & & \\[6pt]
  R3 & & \\[2pt]
  \bottomrule
\end{tabularx}

% ========================= BOM =========================================
\section{Bill of Materials --- max \$50 total (excl.\ Raspberry Pi / Arduino)}

\renewcommand{\arraystretch}{1.22}
\begin{tabularx}{\linewidth}{@{}p{0.24\linewidth} p{0.16\linewidth} p{0.12\linewidth} c c p{0.10\linewidth} X@{}}
  \toprule
  \textbf{Component} & \textbf{Part \# / Model} & \textbf{Supplier} &
  \textbf{Unit \$} & \textbf{Qty} & \textbf{Have /\newline Need} &
  \textbf{Why this part? (link to R1--R3)} \\
  \midrule
  & & & & & & \\[5pt]
  & & & & & & \\[5pt]
  & & & & & & \\[5pt]
  & & & & & & \\[5pt]
  & & & & & & \\[3pt]
  \midrule
  \multicolumn{3}{r}{\textbf{Total ``Need'' items:}} &
  \multicolumn{4}{l}{\underline{\hspace{2.4cm}}} \\
  \bottomrule
\end{tabularx}

\vspace{2pt}
{\small\textit{Datasheet / source for each sensor or module (URL or document name + section):}}\par
\noindent\hrulefill\par
\noindent\hrulefill

% ========================= GRADING ====================================
\vspace{3pt}
\begin{mdframed}[backgroundcolor=notebg, linecolor=noteborder, linewidth=0.8pt,
                 innertopmargin=4pt, innerbottommargin=4pt,
                 innerleftmargin=7pt, innerrightmargin=7pt]
\small
\begin{tabularx}{\linewidth}{@{}p{0.52\linewidth} X@{}}
  \textbf{Grading} & \textbf{Instructor note} \\[2pt]
  BOM with part numbers but no justification: \textbf{50\%} &
  \multirow{3}{*}{Specs due Wed 03/19 end of day.
  Parts ordered Mon 03/23 --- you need them for Week~14.
  If ``Need'' total $>$ \$50, talk to me before submitting.} \\
  BOM + costs + ``Have/Need'' filled: \textbf{80\%} & \\
  Full BOM + PES link + datasheet cited for every part: \textbf{100\%} & \\
\end{tabularx}
\end{mdframed}

\end{document}
